---
title: "Term Project"
author: "Leslie Speight Youtsey and Elizabeth Weatherup"
date: "10/22/2020"
output: html_document
---

```{r}
bats_flux <- read.csv("bats_flux.csv")

delete.na <- function(DF, n=0) {
  DF[rowSums(is.na(DF)) <= n,]
} #Function that takes rows out that contains NAs 

bats_flux.noNA <- delete.na(bats_flux) # Take out rows containing NAs 


Data <- subset(bats_flux,select = c("cr","dep","yymmdd1","yymmdd2","Lat2","Lat2.1","Long1","Long2","M_avg","C_avg","N_avg","P_avg"))

Data.noNA <- delete.na(Data)

#Data_noNA <- write.csv(x= Data.noNA,file = "Data_noNA.csv")
```


ANOVA Trial 
```{r}
#we want to make predictions on the Mass average flux. 

# we want to compare this against depth and C,N,P fluxes 

# Depth is categorical 
# N,P,C are numerical 

library(psych) # for `pairs.panels()`
library(lattice)

ANOCOVA.data <- subset(Data.noNA, select = c("dep", "M_avg", "C_avg","N_avg","P_avg"))
as.character(ANOCOVA.data$dep) # need to make dep categorical not numerical 
attach(ANOCOVA.data)

summary(ANOCOVA.data)

pairs.panels(data.frame(dep,C_avg,N_avg, P_avg, M_avg))

xyplot(M_avg ~ C_avg | dep, 
                  main="Fig2a: Activity level-specific scatterplots"
           )

    xyplot(M_avg ~ C_avg, groups=dep, 
                  auto.key=TRUE,
                  main="Fig2b: Scatterplot with color=group level"
           )
    
    xyplot(M_avg ~ C_avg, groups=dep, 
                  type=c("p","r"), # `p` for _points_, `r` for _regression line_
                                   # see https://stackoverflow.com/questions/12972039/plotting-xyplot-with-regression-line-on-lattice-graphics for more
                  auto.key=TRUE,
                  main="Fig3: Three standalone depth level-specific `lm()` fits"
           )
```

Bar graph trial 
```{r}


```


Looking at fluxes over depths and hypothesizing that flux will decrease over depth 
```{r}
# separate each flux into its own dataframe 
# delete rows that contain NAs 
C_flux.data <- delete.na(subset(Data,select = c("cr","dep","yymmdd1","yymmdd2","Lat2","Lat2.1","Long1","Long2","C_avg")))
N_flux.data <- delete.na(subset(Data,select = c("cr","dep","yymmdd1","yymmdd2","Lat2","Lat2.1","Long1","Long2","N_avg")))
P_flux.data <- delete.na(subset(Data,select = c("cr","dep","yymmdd1","yymmdd2","Lat2","Lat2.1","Long1","Long2","P_avg")))
M_flux.data <- delete.na(subset(Data,select = c("cr","dep","yymmdd1","yymmdd2","Lat2","Lat2.1","Long1","Long2","M_avg")))


# Subset each depth and rename columns 
library(plyr)
C_flux.depth.150 <- subset(C_flux.data, dep == 150)
C_flux.depth.150 <- rename(C_flux.depth.150, c("dep"="dep.150"))
C_flux.depth.150 <- rename(C_flux.depth.150, c("C_avg" ="C_avg.150"))
C_flux.depth.200 <- subset(C_flux.data, dep == 200)
C_flux.depth.200 <- rename(C_flux.depth.200, c("dep"="dep.200"))
C_flux.depth.200 <- rename(C_flux.depth.200, c("C_avg" ="C_avg.200"))
C_flux.depth.300 <- subset(C_flux.data, dep == 300)
C_flux.depth.300 <- rename(C_flux.depth.300, c("dep"="dep.300"))
C_flux.depth.300 <- rename(C_flux.depth.300, c("C_avg" ="C_avg.300"))
C_flux.depth.400 <- subset(C_flux.data, dep == 400)
C_flux.depth.400 <- rename(C_flux.depth.400, c("dep"="dep.400"))
C_flux.depth.400 <- rename(C_flux.depth.400, c("C_avg" ="C_avg.400"))
```


Paired sample t test for each depth and the depth below it 
*This is a paired test because some samples were collected on the same cruise and at the same location. So for each flux value at one depth, there was a flux value collected at a different depth at the same location!
* Data frames were merged based on cruise, year, and location to have the same number of values.  

(1.1)

$\mu_1$ = mean Carbon flux at depth 150
$\mu_2$ = mean Carbon flux at depth 200

$H_0$ : Carbon flux does not change between 150 and 200 meter. $\mu_1$ = $\mu_2$
$H_1$ : Carbon flux is lower at 200 meter then at 150 meter. $\mu_1$ > $\mu_2$

```{r}
C.depth.150.200 <- merge(C_flux.depth.150,C_flux.depth.200,by = c("cr","yymmdd1","yymmdd2","Lat2","Lat2.1","Long1","Long2"))

t.test(C.depth.150.200$C_avg.150 ,C.depth.150.200$C_avg.200, alternative = "greater", paired = TRUE, var.equal = TRUE)

```

(1.2)

$\mu_1$ = mean Carbon flux at depth 200
$\mu_2$ = mean Carbon flux at depth 300

$H_0$ : Carbon flux does not change between 200 and 300 meter. $\mu_1$ = $\mu_2$
$H_1$ : Carbon flux is lower at 300 meter then at 200 meter. $\mu_1$ > $\mu_2$

```{r}
C.depth.200.300 <- merge(C_flux.depth.200,C_flux.depth.300,by = c("cr","yymmdd1","yymmdd2","Lat2","Lat2.1","Long1","Long2"))

t.test(C.depth.200.300$C_avg.200 ,C.depth.200.300$C_avg.300, alternative = "greater", paired = TRUE, var.equal = TRUE)
```

(1.3)

$\mu_1$ = mean Carbon flux at depth 300
$\mu_2$ = mean Carbon flux at depth 400

$H_0$ : Carbon flux does not change between 300 and 400 meter. $\mu_1$ = $\mu_2$
$H_1$ : Carbon flux is lower at 400 meter then at 300 meter. $\mu_1$ > $\mu_2$

```{r}
C.depth.300.400 <- merge(C_flux.depth.300,C_flux.depth.400,by = c("cr","yymmdd1","yymmdd2","Lat2","Lat2.1","Long1","Long2"))

t.test(C.depth.300.400$C_avg.300 ,C.depth.300.400$C_avg.400, alternative = "greater", paired = TRUE, var.equal = TRUE)
```






# Title : Analyze water quality data 

# Abstract: So this project will involve bla bla bla 

# Introduction: 

# Toolbox discussion:

# Description:

# Implementation:

# Results discussion:

# Limitations:

# References:

